---
title: "Ecoli_charging"
author: "Chris Katanski"
date: "3/15/2021"
output: html_document
---

#Set up packages
```{r, set up packages etc., echo=FALSE, results="hide", warning=FALSE, message=FALSE}
#Improt packages for plotting and data manipulation.
library(ggplot2)
library(scales)
library(reshape2)
library(tidyr)
library(dplyr) #its important that this is after tidyr I think

library(grid)
library(gridExtra)
library(ggpubr)
library(errors)

#Reading in svg files into r for ggplotting
library(grImport2)
library(grConvert)
#Reading and plotting jpeg and PNG
library(jpeg)
library(png)
library(svglite)

library(ggrepel)
library(stringr)
library(forcats)
library(readxl)

#For nice log labels
#library(cat.extras)
#library(plotly)

library(extrafont)
font_import()
loadfonts()

#Set the global figure theme now
theme_set(
  theme_bw() + theme(#legend.position = "None", 
        text = element_text(family = "Arial", size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"))
  )

#Define a function that excludes things
`%nin%` = Negate(`%in%`)

empty_graph <- ggplot() + theme_void()
```

#Set your working directory
```{r}
#Probably chance this line for personal use
#Where did you put the data and this R script? Navigate to there
setwd("~/4SR/data/20210315_Ecoli_charging_collaboration/source/")


```

#Read in data
##Base_wise
```{r}
PATH="../5_tsv/Escherichia_tRNA_reference_T/"
FILES <-data.frame( file_name = list.files(PATH) ) %>%
  filter(!grepl("unassigned",file_name))

#Split generic file names into salient parts
#Some of this script is designed for tRNA fragment analysis, but I didn't do fragments yet
#for these samples, so it's slightly redundant for now.
FILES <- FILES %>%
  separate(file_name, sep="_", c("library","junk1", "junk2", "barcode", "junk4", "bin_start", "bin_stop", "junk3"), fill="right", remove=FALSE) %>%
  select(-junk1, -junk2, -junk3, -junk4) 

FILES <- FILES %>%
  mutate(treatment = barcode, #genotype
         rep = barcode, #replicate
         #DM = library #These samples were not DM treated
         ) %>%
  select(-library, -barcode)

#PROJECT SPECIFIC DECODING
FILES$file_name <- as.character(FILES$file_name)
FILES$treatment <- recode(FILES$treatment, 
                          "bc7"="WT",
                          "bc8"="WT",
                          "bc9"="WT",
                          "bc10"="KO",
                          "bc11"="KO",
                          "bc12"="KO"
                          )
FILES$rep <- recode(FILES$rep, 
                    "bc7"="1",
                    "bc8"="2",
                    "bc9"="3",
                    "bc10"="1",
                    "bc11"="2",
                    "bc12"="3"
                          )

#=======================+
#Add a place holder value for "bin_start" if no fragment data is availible
FILES <- FILES %>%
  mutate(bin_start = ifelse(is.na(bin_start), -3, bin_start),
         bin_stop = ifelse(is.na(bin_stop), -3, bin_stop))
#========================


read_in_one <- function(row){
  output <- read.csv(paste0(PATH, row$file_name), header=T, sep="\t") %>%
    mutate(rep = row$rep,
           treatment = row$treatment,
           bin_start = row$bin_start,
           bin_stop  = row$bin_stop,)
  #specify data types since enpty dataframes get confused and sad
  output$gene <- as.character(output$gene)
  output$base <- as.character(output$base)
  return(output)
}

Ecoli_charging_base_data <- FILES %>%
  group_by(file_name) %>%
  do(read_in_one(.)) %>%
  ungroup() %>%
  separate(gene, sep="_", into=c("j1", "j2", "j3", "j4", "j5", "j6", "gene")) %>%
  select(-j1, -j2, -j3, -j4, -j5, -j6) %>%
  separate(gene, sep="-", into=c("j1","AA","anticodon","number"), remove=F) %>%
  select(-j1) %>%
  filter(!is.na(gene))
```


##Counts
```{r}
PATH="../6_sam_counter/Escherichia_tRNA_reference_T/"
FILES <-data.frame( file_name = list.files(PATH) ) %>%
  filter(!grepl("unassigned", file_name),
         file_name %nin% c("kallisto.out"))

#Split generic file names into salient parts
FILES <- FILES %>%
  separate(file_name, sep="_", c("library","junk1", "junk4","barcode", "junk2", "bin_start", "bin_stop", "junk3"), fill="right", remove=FALSE) %>%
  select(-junk1, -junk2, -junk3, -junk4) 

FILES <- FILES %>%
  mutate(treatment = barcode,
         rep = barcode) 

#PROJECT SPECIFIC DECODING
FILES$file_name <- as.character(FILES$file_name)
FILES$treatment <- recode(FILES$treatment, 
                          "bc7"="WT",
                          "bc8"="WT",
                          "bc9"="WT",
                          "bc10"="KO",
                          "bc11"="KO",
                          "bc12"="KO"
                          )
FILES$rep <- recode(FILES$rep, 
                    "bc7"="1",
                    "bc8"="2",
                    "bc9"="3",
                    "bc10"="1",
                    "bc11"="2",
                    "bc12"="3"
                          )

#=======================+
FILES <- FILES %>%
  mutate(bin_start = ifelse(is.na(bin_start), -3, bin_start),
         bin_stop = ifelse(is.na(bin_stop), -3, bin_stop))
#========================

read_in_one <- function(row){
  output <- read.csv(paste0(PATH, row$file_name), header=T, sep="\t") %>%
    mutate(rep = row$rep,
           treatment = row$treatment,
           bin_start = row$bin_start,
           bin_stop  = row$bin_stop,)
  #specify data types since enpty dataframes get confused and sad
  output$name <- as.character(output$name)
  output <- output %>%
    mutate(gene=name) %>%
    select(-name)
  return(output)
}

Ecoli_charging_counts_data <- FILES %>%
  group_by(file_name) %>%
  do(read_in_one(.)) %>%
  ungroup() %>%
  separate(gene, sep="_", into=c("j1", "j2", "j3", "j4", "j5", "j6", "gene")) %>%
  select(-j1, -j2, -j3, -j4, -j5, -j6) %>%
  separate(gene, sep="-", into=c("j1","AA","anticodon","number"), remove=F) %>%
  select(-j1) %>%
  filter(!is.na(gene))
```


#Calculate charging
(PROBABLY FLAWED BECAUSE HAIRPIN OLIGO ENDS WITH "A")
```{r}
#Calculate charging
temp <- Ecoli_charging_base_data %>%
  select(-bin_stop) %>%
  filter(bin_start=="-3") %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  mutate(max_position = max(position)) %>%
  filter(position %in% c(max_position, max_position-1)) %>%
  mutate(charging = ifelse(position==max_position, "end", "penultimate")) %>%
  filter((charging=="end" & base=="A")  |
         (charging=="penultimate" & base=="C")) %>%
  select(treatment, rep, bin_start,  gene, AA, anticodon, number, pileup, charging) %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  pivot_wider(names_from = c("charging"), values_from = c("pileup")) %>%
  mutate(charging= end/penultimate)

#Add a mean charging for each gene
temp <- temp %>%
  ungroup() %>%
  mutate(charging = set_errors(charging) ) %>%
  group_by(treatment, gene) %>%
  mutate(mean_charging = mean(charging, na.rm=T))

#Graph the data
plot1 <- ggplot(filter(temp,
              AA !="Und"),
       aes(x=interaction(AA, anticodon, number, lex.order = T), y=charging, color=treatment)) +
  geom_point() +
  labs(color="Genotype") +
  geom_point(aes(y=mean_charging), shape=95, size=5) +
  theme(axis.text.x = element_text(angle=90, vjust=0.5),
        axis.title.x = element_blank())

#Format the graph so it can be saved easily
layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)
#Set path and name and size of saved graph
path="../figures/"
figure_name="charging_genewise"
width=7
height=3
scale=1.5
dpi=600
#Save the graph
##SVG format
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
##PNG format
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)

```

#Global charging change
```{r}
#Calculate charging
temp <- Ecoli_charging_base_data %>%
  select(-bin_stop) %>%
  filter(bin_start=="-3") %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  mutate(max_position = max(position)) %>%
  filter(position %in% c(max_position, max_position-1)) %>%
  mutate(charging = ifelse(position==max_position, "end", "penultimate")) %>%
  filter((charging=="end" & base=="A")  |
         (charging=="penultimate" & base=="C")) %>%
  select(treatment, rep, bin_start,  gene, AA, anticodon, number, pileup, charging) %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  pivot_wider(names_from = c("charging"), values_from = c("pileup")) %>%
  mutate(charging= end/penultimate)

#For p-value calculation, specificy what to compare (works with stat_compare_means, below)
my_comparisons <- list( c("WT", "KO") )
#Reorder WT and KO so they look nice on the graph
temp$treatment <- factor(temp$treatment, levels=c("WT","KO"))

#Make nice graph with p-value
plot1 <- ggplot(filter(temp,
              AA !="Und"),
       aes(x=treatment, y=charging)) +
  geom_boxplot() +
  labs(color="Genotype") +
  stat_compare_means(comparisons = my_comparisons, size=3,  method="t.test")  +
  theme(#axis.text.x = element_text(angle=90, vjust=0.5),
        axis.title.x = element_blank()
        )
#format the graph for easy saving
layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)
#Specificy name and size etc for figure saving
path="../figures/"
figure_name="charging_global"
width=2
height=2.5
scale=1.5
dpi=600
#Save the figure
##SVG format
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
##PNG format
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)


```

##Ratio change in charging?
```{r}
#Not quite sure what Tao wants here, but I'll try and interpret something
#Calculate charging
temp <- Ecoli_charging_base_data %>%
  select(-bin_stop) %>%
  filter(bin_start=="-3") %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  mutate(max_position = max(position)) %>%
  filter(position %in% c(max_position, max_position-1)) %>%
  mutate(charging = ifelse(position==max_position, "end", "penultimate")) %>%
  filter((charging=="end" & base=="A")  |
         (charging=="penultimate" & base=="C")) %>%
  select(treatment, rep, bin_start,  gene, AA, anticodon, number, pileup, charging) %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  pivot_wider(names_from = c("charging"), values_from = c("pileup")) %>%
  mutate(charging= end/penultimate)

#Take ratio of KO to WT
temp <- temp %>%
  group_by(treatment, gene, AA, anticodon, number) %>%
  summarise(mean_charging = mean(charging, na.rm=T)) %>%
  ungroup() %>%
  select(treatment, mean_charging, gene, AA, anticodon, number) %>%
  pivot_wider(names_from=c("treatment"), values_from=c("mean_charging"))  %>%
  mutate(charging_ratio = KO / WT)

ggplot(filter(temp,
              AA!="Und"),
       aes(x=charging_ratio)) +
  geom_density()
#I don't think this is a particularly insightful analysis

```




#Abundance changes
```{r}
#convert to rpm
temp <- Ecoli_charging_counts_data %>%
  group_by(rep, treatment) %>%
  mutate(rpm = count / sum(count)) %>%
  ungroup %>%
  mutate(rpm = set_errors(rpm)) %>%
  group_by(gene, treatment) %>%
  mutate(mean_rpm = mean(rpm, na.rm=T))


plot1 <- ggplot(filter(temp,
              AA !="Und"),
       aes(x=interaction(AA, anticodon, number, lex.order = T), y=rpm, color=treatment)) +
  geom_point() +
  labs(color="Genotype") +
  geom_point(aes(y=mean_rpm), shape=95, size=5) +
  theme(axis.text.x = element_text(angle=90, vjust=0.5),
        axis.title.x = element_blank())

layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)

path="../figures/"
figure_name="abundance_genewise"
width=7
height=3
scale=1.5
dpi=600
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)

```


#Global abundance change
```{r}
#convert to rpm
temp <- Ecoli_charging_counts_data %>%
  group_by(rep, treatment) %>%
  mutate(rpm = count / sum(count)) %>%
  ungroup %>%
  mutate(rpm = set_errors(rpm)) %>%
  group_by(gene, treatment, AA, anticodon, number) %>%
  summarise(mean_rpm = mean(rpm, na.rm=T))


#For p-value calculation, specificy what to compare (works with stat_compare_means, below)
my_comparisons <- list( c("WT", "KO") )
#Reorder WT and KO so they look nice on the graph
temp$treatment <- factor(temp$treatment, levels=c("WT","KO"))

#Make nice graph with p-value
plot1 <- ggplot(filter(temp,
              AA !="Und"),
       aes(x=treatment, y=mean_rpm)) +
  geom_boxplot() +
  labs(color="Genotype") +
  stat_compare_means(comparisons = my_comparisons, size=3,  method="t.test")  +
  theme(#axis.text.x = element_text(angle=90, vjust=0.5),
        axis.title.x = element_blank()
        )


layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)

path="../figures/"
figure_name="abundance_global"
width=2
height=2.5
scale=1.5
dpi=600
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)

```


#Global abundance ratio
```{r}
#convert to rpm
temp <- Ecoli_charging_counts_data %>%
  group_by(rep, treatment) %>%
  mutate(rpm = count / sum(count)) %>%
  ungroup %>%
  mutate(rpm = set_errors(rpm)) %>%
  group_by(gene, treatment, AA, anticodon, number) %>%
  summarise(mean_rpm = mean(rpm, na.rm=T))

#Take ratio to WT genewise
temp <- temp %>%
  ungroup() %>%
  filter(treatment=="WT") %>%
  mutate(WT_mean_rpm = mean_rpm) %>%
  select(gene, WT_mean_rpm) %>%
  full_join(temp, ., by=c("gene")) %>%
  filter(treatment == "KO")

plot1 <- ggplot(filter(temp,
              AA !="Und"),
       aes(x=interaction(AA, anticodon, number, lex.order = T), y=mean_rpm / WT_mean_rpm)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype=2, alpha=0.3) +
  theme(axis.text.x = element_text(angle=90, vjust=0.5),
        axis.title.x = element_blank())

layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)

path="../figures/"
figure_name="abundance_genewise_ratio"
width=7
height=3
scale=1.5
dpi=600
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)



```

#Bizzare abundance boxplot
```{r}
#convert to rpm
temp <- Ecoli_charging_counts_data %>%
  group_by(rep, treatment) %>%
  mutate(rpm = count / sum(count)) %>%
  ungroup %>%
  mutate(rpm = set_errors(rpm)) %>%
  group_by(gene, treatment, AA, anticodon, number) %>%
  summarise(mean_rpm = mean(rpm, na.rm=T))

temp <- temp %>%
  select(gene, treatment, mean_rpm, AA, anticodon, number) %>%
  pivot_wider(names_from=c("treatment"), values_from=c("mean_rpm")) %>%
  mutate(KO_WT_ratio = KO / WT)

plot1 <- ggplot(filter(temp, AA !="Und"),
       aes(y=KO_WT_ratio)) +
  geom_boxplot() +
  geom_text_repel(data=filter(temp, KO_WT_ratio>1.25 | KO_WT_ratio <0.65),
    aes(label=paste0(AA, anticodon, number)), x=0) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)

path="../figures/"
figure_name="abundance_global_ratio"
width=2
height=2.5
scale=1.5
dpi=600
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)

```

#Modifications
```{r}
#Average the modifcaiton rate across replicates
temp <- Ecoli_charging_base_data %>%
  ungroup() %>%
  #mutate(mutation = set_errors(mutation)) %>% #Good for error propegation, but not necessary here
  group_by(treatment, position, gene, AA, anticodon, number) %>%
  summarise(mutation = mean(mutation, na.rm=T)) %>%
  mutate(mutation = as.numeric(mutation))

temp$treatment <- factor(temp$treatment, level=c("WT","KO"))

plot1 <- ggplot(filter(temp,
              AA!="Und",
              #AA=="Arg"
              ),
       aes(x=position, y=interaction(AA, anticodon, number, lex.order = T), fill=mutation)) +
  geom_tile() +
  #very fancy coloring
  scale_fill_gradient2(low="blue", mid="grey70", midpoint=0, high="red", na.value="white",
                       limits=c(0, 0.3), breaks=c(0,0.1,0.2,0.3)
                       #limits=c(0, .5), breaks=c(0,0.25,0.5)
                       )+
  #extremely fancy coloring
  geom_tile(data=filter(temp, AA!="Und", mutation >=0.3), fill="red4") +
  facet_grid(cols=vars(treatment)) +
  theme(strip.background = element_blank(),
        axis.title.y = element_blank())


#format the graph for easy saving
layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)
#Specificy name and size etc for figure saving
path="../figures/"
figure_name="mutation_rate"
width=7
height=4
scale=1.75
dpi=600
#Save the figure
##SVG format
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
##PNG format
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)

```


#Change in modifications
```{r}
temp <- Ecoli_charging_base_data %>%
  ungroup() %>%
  #mutate(mutation = set_errors(mutation)) %>%
  group_by(treatment, position, gene, AA, anticodon, number) %>%
  summarise(mutation = mean(mutation, na.rm=T)) %>%
  pivot_wider(names_from = c("treatment"), values_from=c("mutation")) %>%
  mutate(delta_mutation = KO - WT) %>%
  mutate(delta_mutation = as.numeric(delta_mutation))

temp$treatment <- factor(temp$treatment, level=c("WT","KO"))

plot1 <- ggplot(filter(temp,
              AA!="Und",
              #AA=="Arg"
              ),
       aes(x=position, y=interaction(AA, anticodon, number, lex.order = T), fill=delta_mutation)) +
  geom_tile() +
  #very fancy coloring
  scale_fill_gradient2(low="blue", mid="grey70", midpoint=0, high="red", na.value="white",
                       limits=c(-0.1, 0.1), breaks=c(-0.1, 0, 0.1)
                       )+
  labs(fill="Change in mutation\n(KO minus WT)") +
  #extremely fancy coloring
  theme(strip.background = element_blank(),
        axis.title.y = element_blank())


#format the graph for easy saving
layout <- rbind(c(1))
figure <- grid.arrange(plot1,  layout_matrix = layout)
#Specificy name and size etc for figure saving
path="../figures/"
figure_name="mutation_change"
width=4
height=4
scale=1.75
dpi=600
#Save the figure
##SVG format
# ggsave(paste0(path, figure_name,".svg"),
#        plot = figure,
#        scale = scale,
#        dpi = dpi,
#        width = width,
#        height = height)
##PNG format
ggsave(paste0(path, figure_name,".png"),
       plot = figure,
       scale = scale,
       dpi = dpi,
       width = width,
       height = height)


```






#_
#2023 03 29
##Table of values
```{r}

#Calculate charging
temp <- Ecoli_charging_base_data %>%
  select(-bin_stop) %>%
  filter(bin_start=="-3") %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  mutate(max_position = max(position)) %>%
  filter(position %in% c(max_position, max_position-1)) %>%
  mutate(charging = ifelse(position==max_position, "end", "penultimate")) %>%
  filter((charging=="end" & base=="A")  |
         (charging=="penultimate" & base=="C")) %>%
  select(treatment, rep, bin_start,  gene, AA, anticodon, number, pileup, charging) %>%
  group_by(treatment, rep, bin_start, gene, AA, anticodon, number) %>%
  pivot_wider(names_from = c("charging"), values_from = c("pileup")) %>%
  mutate(charging= end/penultimate) %>%
  ungroup() %>%
  select(-bin_start, -penultimate, -end)  %>%
  pivot_wider(names_from = c(treatment, rep),  values_from = c(charging) )

write.csv(temp, file="./Double_KO_charging.csv", row.names = F)

```



##Abundance table of values
```{r}

#convert to rpm
temp <- Ecoli_charging_counts_data %>%
  group_by(rep, treatment) %>%
  mutate(rpm = count / sum(count)) %>%
  # ungroup %>%
  # mutate(rpm = set_errors(rpm)) %>%
  # group_by(gene, treatment) %>%
  # mutate(mean_rpm = mean(rpm, na.rm=T)) %>%
  ungroup() %>%
  select(-bin_start, -bin_stop, -count, -file_name) %>%
  pivot_wider(names_from = c(treatment, rep),  values_from = c(rpm) )

write.csv(temp, file="./Double_KO_abundance.csv", row.names = F)

```
